---
title: "Introduction to OmniAgeR"
author:
- name: Zhaozhen Du
  affiliation: Shanghai Institute of Nutrition and Health, CAS
  email: duzhaozhen2022@sinh.ac.cn
- name: Andrew E. Teschendorff
  affiliation: Shanghai Institute of Nutrition and Health, CAS
  email: andrew@sinh.ac.cn
package: OmniAgeR
output: 
  BiocStyle::html_document:
    number_sections: true
    toc: yes
vignette: |
  %\VignetteIndexEntry{Introduction to OmniAgeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview of Implemented Clocks

The `OmniAgeR` package provides two primary, high-level interfaces: epiMarker():
For calculating a comprehensive suite of aging-related clocks (chronological, 
biological, mitotic, etc.). For advanced users or specific applications, the 
package also provides direct access to individual clock functions (e.g., 
horvath2013Clock()) and specialized "bundle" calculators (e.g., pcClocks(),
systemsAge()) that manage complex dependencies. These tools are broadly 
categorized into three main groups:

1.  **Epigenetic (DNAm) Clocks**
2.  **Transcriptomic (RNA) Clocks**
3.  **Other Aging-Related Surrogate Biomarkers**

The suite of **epigenetic aging clocks** is particularly extensive, organized 
into several functional categories, including predictors for chronological age, 
biological age, cellular division (mitotic clocks), and gestational age.

# Epigenetic (DNAm) Aging Clocks

## Chronological Age Clocks

These DNAm clocks are designed for the prediction of chronological age.

-   `Horvath2013`: (Horvath 2013) The original pan-tissue clock from 353 CpGs.
-   `Hannum`: (Hannum et al. 2013) Blood-specific clock from 71 CpGs.
-   `Lin`: (Lin et al. 2016) Clock based on 99 CpGs.
-   `VidalBralo`: (Vidal-Bralo et al. 2016) Clock based on 8 CpGs.
-   `ZhangClock`: (Zhang et al. 2019) Improved-precision clock from 514 CpGs.
-   `Horvath2018`: (Horvath et al. 2018) Skin & blood clock.
-   `Bernabeu_cAge`: (Bernabeu et al. 2022) Clock based on 3225 CpGs.
-   `CorticalClock`: (Shireby et al. 2020) Brain cortical clock.
-   `PedBE`: (McEwen et al. 2019) Pediatric buccal epithelial clock for children.
-   `CentenarianClock`: (Eric Dec et al. 2023) Centenarian epigenetic clocks.
-   `Retro_age`: (Ndhlovu et al. 2024) Retroelement-based clock developed on 
    the EPIC v1/v2 arrays.
-   `ABEC`: (Lee et al. 2020) Adult blood-based EPIC clock.
-   `eABEC`: (Lee et al. 2020) Extended adult blood-based EPIC clock.
-   `cABEC`: (Lee et al. 2020) Common adult blood-based EPIC clock.
-   `PipekElasticNet`: (Pipek et al. 2023) Pipek's multi-tissue elastic Net 
    epigenetic clock (239 CpGs).
-   `PipekFilteredh`: (Pipek et al. 2023) Pipek's filtered horvath epigenetic 
    clock (272 CpGs).
-   `PipekRetrainedh`: (Pipek et al. 2023) Pipek's retrained horvath epigenetic
    clock (308 CpGs).
-   `WuClock`: (Wu et al. 2019) Wu's Epigenetic Clock for Pediatric age 
    estimation.
-   `Weidner`: (Weidner et al. 2014) Calculate weidner epigenetic age (3 CpGs).
-   `IntrinClock`: (Tomusiak et al. 2024) Calculates the intrinsic cellular age.
-   `Garagnani`: (Garagnani et al. 2012) The Garagnani ELOVL2-based epigenetic 
    age score(1 CpG).
-   `PCHorvath2013`, `PCHorvath2018`, `PCHannum`: (Higgins-Chen et al. 2022) 
    Computationally-bolstered versions of the original clocks. These PC clocks 
    are retrained using principal components (PCs) derived from a large CpG set 
    to minimize technical noise and improve reliability

## Biological Age Clocks

These advanced clocks are designed to capture biological aging rather than 
chronological time, often showing stronger associations with health outcomes 
and mortality.

-   `Zhang10`: (Zhang et al. 2017) 10-CpG clock associated with mortality.
-   `PhenoAge`: (Levine et al. 2018) Predicts phenotypic age from 513 CpGs.
-   `DunedinPACE`: (Belsky et al. 2022) Quantifies the pace of biological aging.
-   `GrimAge1`: The original GrimAge clock (Lu et al. 2019).
-   `GrimAge2`: (Lu et al. 2022) Updated composite biomarker of mortality risk.
-   `PCPhenoAge`, `PCGrimAge1`: (Higgins-Chen et al. 2022) 
    Computationally-bolstered PC versions of the PhenoAge and GrimAge1 clocks,
    retrained on principal components for enhanced reliability.
-   `DNAmFitAge`: (McGreevy et al. 2023) Biological age indicator incorporating
    DNAmGrimAge and 3 DNAm-based physical fitness markers.
-   `IC_Clock`: (Fuentealba et al. 2025) The intrinsic capacity (IC) clock based
    on 91 CpGs.
-   `SystemsAge`: (Sehgal et al. 2025) The systems age and 11 system-specific 
    scores.

## Cellular Aging Clocks

This category groups biomarkers that measure two key forms of cellular aging: 
cell proliferation (mitotic clocks) and telomere attrition (DNAmTL).

### Mitotic Clocks (Cellular Division)

A specialized set of clocks that measure cell division history or stem cell 
divisions.

-   `epiTOC1`: (Yang et al. 2016) Average beta value of 385 promoter CpGs.
-   `epiTOC2`: (Teschendorff et al. 2020) Estimates stem cell divisions using a 
    dynamic model.
-   `epiTOC3`: Estimates stem cell divisions based on unmethylated population 
    doubling associated CpGs.
-   `stemTOCvitro`: (Zhu et al. 2024) The 0.95 upper quantile of 629 
    stemTOCvitro CpGs (promoter CpGs unmethylated in fetal tissue that 
    hypermethylate with population-doublings).
-   `stemTOC`: (Zhu et al. 2024) The 0.95 upper quantile of 371 stemTOC CpGs, 
    filtered for in-vivo hypermethylation with age.
-   `RepliTali`: (Endicott et al. 2022) Based on 87 population doubling 
    associated hypomethylated CpGs.
-   `HypoClock`: (Teschendorff et al. 2020) Based on hypomethylation at 678 
    solo-WCGW sites.
-   `EpiCMIT_hyper`: (Duran-Ferrer et al. 2020) Average beta value of 184 
    age-associated hypermethylated CpGs.
-   `EpiCMIT_hypo`: (Duran-Ferrer et al. 2020) Average beta value of 1164 
    age-associated hypomethylated CpGs.

### DNAm Telomere Length (TL) Clocks

-   `DNAmTL`: (Lu AT et al. 2019) Calculates the Leukocyte telomere length.
-   `PCDNAmTL`: (Lu AT et al. 2019; Higgins-Chen et al. 2022) 
    A computationally-bolstered "PC" version of the original DNAmTL clock. This
    clock is retrained using principal components (PCs) derived from a large CpG
    set to minimize technical noise and improve reliability.

## Causal Clocks

A new generation of clocks developed to reflect potential causal drivers 
of aging.

-   `CausalAge`, `DamAge`, `AdaptAge`: (Ying et al. 2024) Three clocks derived 
from a causality-enriched model. They dissect aging into distinct components: 
'Causal' (CausalAge), 'Damage' (DamAge), and Adaptation' (AdaptAge).

## Stochastic Clocks

-   `StocH`, `StocP`, `StocZ`: (Tong et al. 2024) Stochastic analogues of the 
Horvath, PhenoAge, and Zhang clocks, trained on artificial cohorts to quantify 
the stochastic component of epigenetic aging.

## Cell-Type Specific Clocks

These clocks are designed to measure aging in specific cell types or tissues, 
accounting for cellular heterogeneity.

-   `Neu-In`, `Glia-In`: (Tong et al. 2024) Intrinsic clocks for neurons, 
and glia.
-   `Neu-Sin`, `Glia-Sin`, `Hep`: (Tong et al. 2024) Semi-intrinsic clocks for 
neurons, glia, and hepatocytes

## Gestational Age Clocks

This collection includes DNAm clocks designed to predict gestational age 
at birth.

-   `BohlinGA`: (Bohlin et al. 2016) The Bohlin Gestational Age.
-   `EPICGA`: (Haftorn et al. 2021) The EPIC Gestational Age.
-   `LeeGA`: (Lee et al. 2019) A list (LeeControl, LeeRobust, LeeRefinedRobust).
-   `KnightGA`: (Knight et al. 2016) The Knight Gestational Age.
-   `MayneGA`: (Mayne et al. 2017) The Mayne Placental Gestational Age.

## Cross-Species Support
-   `EnsembleAge`:  (Haghani et al. 2025) The multi-clock framework 
    implementation. This includes mouse-specific estimators 
    (`EnsembleAgeMouse_Dynamic` and `EnsembleAgeMouse_Static`) and a 
    cross-species (`EnsembleAgeMouse_HumanMouse`) predictor applicable to both 
    humans and mice

-   `UniversalPanMammalianClocks`: (Lu et al. 2023) The three pan-tissue clocks
    (Clock 1, 2, 3) applicable across all mammalian tissues.
-   `PanMammalianBlood`: (Lu et al. 2023) The two blood-specific pan-mammalian 
    clocks (Clock 2, 3).
-   `PanMammalianSkin`: (Lu et al. 2023) The two skin-specific pan-mammalian 
    clocks (Clock 2, 3).

# Transcriptomic Clocks

In addition to DNAm-based models, the package implements cutting-edge 
transcriptomic (RNA-seq) clocks:

-   `sc-ImmuAging`: A single-cell, immune-cell-type-specific aging clock.
-   `Brain_CT_clock`: A brain-cell-type-specific aging clock based on 
    Single-Nuclei Transcriptomics.
-   `PASTA`: A robust, multi-tissue transcriptomic clock based on a novel 
    age-shift learning strategy. It utilizes rank-normalization to ensure broad 
    applicability across heterogeneous datasets, including bulk RNA-seq, 
    single-cell pseudobulks, and microarrays.

# Surrogate Biomarkers & Scores

Besides, `OmniAgeR` includes functions to compute several other important 
surrogate markers associated with aging, lifestyle, and health status from DNAm 
data.

-   `CRP`: Calculates a score for C-reactive protein (CRP) levels, a marker of 
inflammation.
-   `CHIP`: Calculates scores for Clonal Hematopoiesis of Indeterminate 
Potential (CHIP).
-   `IL6`: Computes a DNA methylation (DNAm) surrogate score for Interleukin-6 
(IL-6) protein levels.
-   `EpiScores`: Computes the 109 validated epigenetic scores (EpiScores) that 
serve as DNA methylation-based proxies for the levels of circulating plasma 
proteins as defined by Gadd et al. (2022).

-   `McCartneyTrait`: Computes epigenetic surrogate scores for 10 different 
complex traits and biomarkers based on blood methylation models. Traits include:
`BMI`, `Smoking` (pack-years), `Alcohol` (units/week), `Education` (years), 
`Total_cholesterol`, `HDL_cholesterol`, `LDL_cholesterol`, `Total_HDL_ratio`, 
`WHR` (Waist-to-Hip Ratio), and `Body_fat_Perc`.

# Disease Risk Prediction

`OmniAgeR` provides a suite of tools for disease risk assessment, including a 
DNA methylation-based smoking index derived from 1,501 CpG sites, as well as 
specialized scores for cancer risks. These tools leverage robust epigenetic 
signatures to quantify environmental exposures and clinical outcomes.

-   `CompSmokeIndex`: Computes a DNAm-based Smoking Index that correlates with 
cancer risk across various tissues.
-   `HepatoXuRisk`: Calculate HepatoXu ctDNA Methylation Scores for 
Hepatocellular Carcinoma


# DNA Methylation Cell-Type Fraction

The OmniAgeR package also provides a DNA Methylation Cell-Type Fraction (CTF) 
clock (implemented as `dnamCTFClock`), which was trained on a large-scale NSPT 
dataset comprising more than 3,500 blood samples. The clock incorporates 12 
immune cell types, estimated using the EpiDISH algorithm, to model aging based 
on cell-type composition.

# Available Clocks and Functions

## Epigenetic (DNAm) Aging Clocks

| Clock Name | Category | Calculator / Function |
|:--------------------|:-----------------|:--------------------------------|
| `Horvath2013` | Chronological | `epiMarker()`,`horvath2013Clock()` |
| `Hannum` | Chronological | `epiMarker()`,`hannumClock()` |
| `Lin` | Chronological | `epiMarker()`,`linClock()` |
| `VidalBralo` | Chronological | `epiMarker()`,`vidalBraloClock()` |
| `ZhangClock` | Chronological | `epiMarker()`,`zhangClock()` |
| `Horvath2018` | Chronological | `epiMarker()`,`horvath2018Clock()` |
| `Bernabeu_cAge` | Chronological | `epiMarker()`,`bernabeuCAge()` |
| `CorticalClock` | Chronological | `epiMarker()`,`corticalClock()` |
| `PedBE` | Chronological | `epiMarker()`,`pedBEClock()` |
| `CentenarianClock` | Chronological | `epiMarker()`,`centenarianClock()` |
| `Retro_age` | Chronological | `epiMarker()`,`retroAge()` |
| `ABEC` | Chronological | `epiMarker()`,`leeABEC()` |
| `eABEC` | Chronological | `epiMarker()`,`leeExtendedABEC()` |
| `cABEC` | Chronological | `epiMarker()`,`leeCommonABEC()` |
| `PipekElasticNet` | Chronological | `epiMarker()`,`pipekElasticNet()` |
| `PipekFilteredh` | Chronological | `epiMarker()`,`pipekFilteredh()` |
| `PipekRetrainedh` | Chronological | `epiMarker()`,`pipekRetrainedh()` |
| `WuClock` | Chronological | `epiMarker()`,`wuClock()` |
| `Weidner` | Chronological | `epiMarker()`,`weidnerClock()` |
| `IntrinClock` | Chronological | `epiMarker()`,`intrinClock()` |
| `Garagnani` | Chronological | `epiMarker()`,`garagnaniClock()` |
| `Zhang10` | Biological | `epiMarker()`,`zhang10()` |
| `PhenoAge` | Biological | `epiMarker()`,`phenoAge()` |
| `DunedinPACE` | Biological | `epiMarker()`,`dunedinPACE()` |
| `GrimAge1` | Biological | `epiMarker()`,`grimAge1()` |
| `GrimAge2` | Biological | `epiMarker()`,`grimAge2()` |
| `DNAmFitAge` | Biological | `epiMarker()`,`dnamFitAge()` |
| `IC_Clock` | Biological | `epiMarker()`,`icClock()` |
| `SystemsAge` | Biological | `epiMarker()`,`systemsAge()` |
| `epiTOC1` | Mitotic | `epiMarker()`,`epiTOC1()` |
| `epiTOC2` | Mitotic | `epiMarker()`,`epiTOC2()` |
| `epiTOC3` | Mitotic | `epiMarker()`,`epiTOC3()` |
| `stemTOC` | Mitotic | `epiMarker()`,`stemTOC()` |
| `stemTOCvitro` | Mitotic | `epiMarker()`,`stemTOCvitro()` |
| `RepliTali` | Mitotic | `epiMarker()`,`repliTali()` |
| `HypoClock` | Mitotic | `epiMarker()`,`hypoClock()` |
| `EpiCMIT_hyper`, `EpiCMIT_hypo` | Mitotic | `epiMarker()`,`epiCMIT()` |
| `DNAmTL` | Telomere length | `epiMarker()`,`dnamTL()` |
| `PCHorvath2013`,`PCHorvath2018`, `PCHannum`, `PCPhenoAge`, `PCGrimAge1`, `PCDNAmTL` | Chronological/Biological/Telomere length | `epiMarker()`,`pcClocks()` |
| `CausalAge`,`DamAge`,`AdaptAge` | Causal | `epiMarker()`,`causalClock()` |
| `StocH`,`StocP`,`StocZ` | Causal | `epiMarker()`,`stochClocks()` |
| `Neu-In`,`Glia-In`,`Neu-Sin`,`Glia-Sin`,`Hep` | Cell-Type Specific | `epiMarker()`,`ctsClocks()` |
| `BohlinGA` | Gestational | `epiMarker()`,`bohlinGa()` |
| `EPICGA` | Gestational | `epiMarker()`,`epicGa()` |
| `LeeGA` | Gestational | `epiMarker()`,`LeeGa()` |
| `KnightGA` | Gestational | `epiMarker()`,`knightGa()` |
| `MayneGA` | Gestational | `epiMarker()`,`mayneGa()` |
| `EnsembleAge_HumanMouse` | Mouse & Human | `epiMarker()`,`ensembleAge()` |
| `EnsembleAge_Static` | Mouse & Human | `epiMarker()`,`ensembleAge()` |
| `EnsembleAge_Dynamic` | Mouse & Human | `epiMarker()`,`ensembleAge()` |
| `UniversalPanMammalianClocks` | PanMammalian | `epiMarker()`,`universalPanMammalianClocks()` |
| `PanMammalianBlood` | PanMammalian | `epiMarker()`,`panMammalianBlood()` |
| `PanMammalianSkin` | PanMammalian | `epiMarker()`,`panMammalianSkin()` |
| `DNAm_CTF_Clock` | Cell-Type Fraction Clock | `epiMarker()`,`dnamCTFClock()` |


## Transcriptomic Clocks

| Clock Name       | Category                  | Calculator / Function |
|:-----------------|:--------------------------|:----------------------|
| `scImmuAging`   | Immune-cell-type-specific  | `scImmuAging()`       |
| `Brain_CT_Clock` | Brain-cell-type-specific  | `brainCtClock()`      |
| `PASTA`          | Multi-tissue              | `pastaScores()`       |

## Surrogate Biomarkers & Scores

| Clock Name | Target | Calculator / Function |
|:--------------------|:-----------------|:--------------------------------|
| `CRP` | CRP/intCRP score | `epiMarker()`, `compCRP()` |
| `CHIP` | CHIP score | `epiMarker()`, `compCHIP()` |
| `IL6` | IL6 score | `epiMarker()`, `compIL6()` |
| `EpiScores` | 109 validated epigenetic scores | `epiMarker()`, `compEpiScores()` |
| `CompSmokeIndex` | Smoking Index | `epiMarker()`, `compSmokeIndex()` |
| `HepatoXuRisk` | Hepatocellular Carcinoma Risk | `epiMarker()`, `hepatoXuRisk()` |
| `SmokeIndex` | Smoking Index | `epiMarker()`, `compSmokeIndex()` |
| `McCartney_Trait` | Epigenetic surrogate scores for 10 different complex traits | `epiMarker()`, `mcCartneyTrait()` |


## Installation

```{r install-bioc, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(c("OmniAgeR", "OmniAgeRData"))
```

During package review or development, the development versions can be installed from GitHub:

```{r install-github, eval=FALSE}
if (!requireNamespace("remotes", quietly = TRUE))
    install.packages("remotes")

remotes::install_github("Duzhaozhen/OmniAgeRData")
remotes::install_github("Duzhaozhen/OmniAgeR")
```

After installation, the packages can be loaded with:

```{r load-packages, message=FALSE}
library(OmniAgeR)
library(OmniAgeRData)
```


# Tutorial Example 1: Apply mitotic clocks on Lung pre-cancerous lesions

## Loading the lung pre-cancerous lesions dataset

This is an Illumina 450k dataset encompassing 21 normal lung tissue samples and 
35 lung carcinoma in situ (LCIS) samples, of which 22 progressed to an invasive 
lung cancer (LC). Here we explore the correlation between mitotic age and cancer 
state with linear model, treating N, LCIS, and LC as ordinal variable (1,2,3) 
and adjusting for age.

```{r mototic-clock-example}
library(OmniAgeR)
library(ggplot2)
library(patchwork)
library(ggpubr)
library(SummarizedExperiment)
lungInv <- loadOmniAgeRdata(
    "omniager_lung_inv",
    verbose = FALSE
)

# Construct the SummarizedExperiment object
lungInvSeObj <- SummarizedExperiment(
    assays = list(beta = lungInv$bmiq_m)
    )

phenoDf <- lungInv$PhenoTypes

my_comparisons <- list(c("N\nN=21", "LCIS\nN=13"), c("LCIS\nN=13", "LCIS->LC\nN=22"))
table(phenoDf$Group)
## Check available epigenetic clocks
listEpiMarker()
```

## Estimation of epigenetic mitotic age

Mitotic age estimation of all the clocks is implemented in a general function *EpiMitClocks*. The required input is a DNAm beta value matrix with rows labeling Illumina 450k/EPIC CpGs and columns labeling samples.

```{r run-mototic-clock}
epiMarkerRes <- suppressPackageStartupMessages(epiMarker(
    x = lungInvSeObj,
    clockNames = "mitotic",
    chronAge = phenoDf$Age,
    minCoverage = 0,
    verbose = FALSE
))
rm(lungInvSeObj)
```

```{r mototic-clock-plot, fig.width=12, fig.height=10, out.width="100%"}
phenoDf$epiTOC1 <- epiMarkerRes$epiTOC1
phenoDf$epiTOC2 <- epiMarkerRes$epiTOC2$irS
phenoDf$epiTOC3 <- epiMarkerRes$epiTOC3$irS
phenoDf$stemTOCvitro <- epiMarkerRes$stemTOCvitro
phenoDf$stemTOC <- epiMarkerRes$stemTOC
phenoDf$HypoClock <- epiMarkerRes$HypoClock
phenoDf$RepliTali <- epiMarkerRes$RepliTali
phenoDf$EpiCMIT_Hyper <- epiMarkerRes$EpiCMIT_Hyper
phenoDf$EpiCMIT_Hypo <- epiMarkerRes$EpiCMIT_Hypo
g <- list()
for (i in 1:9) {
    tempDf <- data.frame("Group" = phenoDf$Group, "Age" = phenoDf$Age, "num" = phenoDf$num, "score" = phenoDf[, i + 3])
    y <- summary(lm(tempDf$score ~ tempDf$num + tempDf$Age))$coefficients[2, 4]
    g[[i]] <- ggplot(tempDf, aes(x = Group, y = score, fill = Group)) +
        guides(fill = "none") +
        geom_boxplot() +
        theme_bw() +
        scale_x_discrete(limits = c("N\nN=21", "LCIS\nN=13", "LCIS->LC\nN=22")) +
        stat_compare_means(method = "wilcox.test", comparisons = my_comparisons, size = 4) +
        xlab("") +
        ylab(colnames(phenoDf)[i + 3]) +
        annotate("text", x = 2, y = max(tempDf$score) * 0.9, label = paste0("P(Age-adjusted)=", signif(y, 2)), size = 4) +
        theme(
            axis.title = element_text(size = 11),
            axis.text = element_text(size = 11)
        )
}

ggpubr::ggarrange(plotlist = g, nrow = 3, ncol = 3)
```

# Tutorial Example 2: Predict chronological age in blood tissue

The epigenetic age of 3 normal blood samples was estimated from their Illumina 
450k methylation profiles using both the Horvath and Zhang epigenetic clocks.

```{r chronological-age-clock}
library(OmniAgeR)
library(ggplot2)
library(patchwork)
library(ggpubr)

data(dnamExample)
hannumBmiqM <- dnamExample[[1]]
phenoTypesHannum <- dnamExample[[2]]
age <- phenoTypesHannum$Age
sex <- ifelse(phenoTypesHannum$Sex == "F", "Female", "Male")
epiMarkerOut <- suppressPackageStartupMessages(epiMarker(hannumBmiqM,
    clockNames = c("Horvath2013", "ZhangClock"),
    minCoverage = 0, verbose = FALSE
))
```

```{r chronological-age-clock-plot, fig.width=8, fig.height=4, out.width="100%"}
gp <- list()

for (i in seq_along(epiMarkerOut)) {
    plot_df <- data.frame(ActualAge = phenoTypesHannum$Age, PredictedAge = epiMarkerOut[[i]])
    cor_test_result <- cor.test(plot_df$ActualAge, plot_df$PredictedAge)
    correlation <- cor_test_result$estimate
    p_value <- cor_test_result$p.value
    mae <- mean(abs(plot_df$PredictedAge - plot_df$ActualAge))
    p_value_formatted <- ifelse(p_value < 0.001,
        formatC(p_value, format = "e", digits = 2),
        round(p_value, 3)
    )
    annotation_text <- paste0(
        "R = ", round(correlation, 3), "\n",
        "P = ", p_value_formatted, "\n",
        "MAE = ", round(mae, 2)
    )
    gp[[i]] <- ggplot(data = plot_df, aes(x = ActualAge, y = PredictedAge)) +
        geom_point(color = "steelblue", size = 3, alpha = 0.8) +
        geom_smooth(method = "lm", color = "black", se = FALSE) +
        annotate("text",
            x = min(plot_df$ActualAge, na.rm = TRUE),
            y = max(plot_df$PredictedAge, na.rm = TRUE),
            label = annotation_text,
            hjust = 0,
            vjust = 1, size = 5
        ) +
        labs(
            title = NULL,
            x = "Chronological Age",
            y = paste0("Predicted Age(", c("Horvath2013", "ZhangClock")[i], ")")
        ) +
        theme_bw() +
        theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5, size = 12),
            axis.title = element_text(size = 14),
            axis.text = element_text(size = 12)
        )
}

ggpubr::ggarrange(plotlist = gp, nrow = 1, ncol = 2)
```

# Tutorial Example 3: Apply sc-ImmuAging to PBMC scRNA-seq dataset

In this example, we showcase the application of the cell-type-specific clock, scImmuAging, to a PBMC scRNA-seq dataset. The dataset includes 5 samples, profiled to contain only CD4+ T cells.

```{r scImmuAging, eval = FALSE}
library(OmniAgeR))
library(glmnet)
library(ggplot2)
library(patchwork)
library(ggpubr)
data("ScPbmcExample")
set.seed(42)
scImmuAgingOut <- suppressPackageStartupMessages(scImmuAging(
   x = ScPbmcExample,
   cellTypes = c("CD4T"),
   assayName = "logcounts",
   donorCol = "donor_id",
   ageCol = "age",
   cellTypeCol = "celltype",
   verbose = FALSE
))
```

```{r scImmuAging-plot, eval = FALSE, fig.width=8, fig.height=4, out.width="100%"}
sc_gp <- list()

for (i in seq_along(scImmuAgingOut)) {
    plot_df <- data.frame(ActualAge = scImmuAgingOut[[i]]$donor$age, PredictedAge = scImmuAgingOut[[i]]$donor$predicted)
    cor_test_result <- cor.test(plot_df$ActualAge, plot_df$PredictedAge)
    correlation <- cor_test_result$estimate
    p_value <- cor_test_result$p.value
    mae <- mean(abs(plot_df$PredictedAge - plot_df$ActualAge))
    p_value_formatted <- ifelse(p_value < 0.001,
        formatC(p_value, format = "e", digits = 2),
        round(p_value, 3)
    )
    annotation_text <- paste0(
        "R = ", round(correlation, 3), "\n",
        "P = ", p_value_formatted, "\n",
        "MAE = ", round(mae, 2)
    )
    sc_gp[[i]] <- ggplot(data = plot_df, aes(x = ActualAge, y = PredictedAge)) +
        geom_point(color = "steelblue", size = 3, alpha = 0.8) +
        geom_smooth(method = "lm", color = "black", se = FALSE) +
        annotate("text",
            x = min(plot_df$ActualAge, na.rm = TRUE),
            y = max(plot_df$PredictedAge, na.rm = TRUE),
            label = annotation_text,
            hjust = 0,
            vjust = 1,
            size = 5
        ) +
        labs(
            title = c("CD4T", "CD8T")[i],
            x = "Chronological Age",
            y = "Predicted Age(sc-ImmuAging)"
        ) +
        theme_bw() +
        theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5, size = 12),
            axis.title = element_text(size = 14),
            axis.text = element_text(size = 12)
        )
}

ggpubr::ggarrange(plotlist = sc_gp, nrow = 1)
```

# Tutorial Example 4: Apply Brain_CT_clock to Brain snRNA-seq dataset

In this example, we showcase the application of the brain cell-type-specific aging Clocks to a brain snRNA-seq dataset. The dataset includes 5 donors, profiled to contain only Oligodendrocytes.

```{r brain-ct-clock}
suppressPackageStartupMessages(library(OmniAgeR))
library(glmnet)
library(ggplot2)
library(patchwork)
library(ggpubr)
library(dplyr)
data("ScBrainExample")

# Recommended Bioconductor workflow: SingleCellExperiment input
clockResults <- suppressPackageStartupMessages(brainCtClock(
    x = ScBrainExample,
    cellTypes = "Oligodendrocytes",
    modelNames = "SC",
    assayName = "logcounts",
    donorCol = "donor_id",
    ageCol = "age",
    cellTypeCol = "celltype",
    verbose = FALSE
))

# Use lapply to iterate over each data frame in the list
averagePredictionsDonor <- lapply(clockResults, function(df) {
    df %>%
        group_by(donorId, age, sampleType, celltype) %>%
        summarise(mean_prediction = mean(prediction, na.rm = TRUE))
})
```

```{r  brain-ct-clock-plot, fig.width=8, fig.height=3, out.width="100%"}
brainPlot <- list()

for (i in seq_along(averagePredictionsDonor)) {
    plot_df <- data.frame(
        ActualAge = averagePredictionsDonor[[i]]$age,
        PredictedAge = averagePredictionsDonor[[i]]$mean_prediction
    )
    cor_test_result <- cor.test(plot_df$ActualAge, plot_df$PredictedAge)
    correlation <- cor_test_result$estimate
    p_value <- cor_test_result$p.value
    mae <- mean(abs(plot_df$PredictedAge - plot_df$ActualAge))
    p_value_formatted <- ifelse(p_value < 0.001,
        formatC(p_value, format = "e", digits = 2),
        round(p_value, 3)
    )
    annotation_text <- paste0(
        "R = ", round(correlation, 3), "\n",
        "P = ", p_value_formatted, "\n",
        "MAE = ", round(mae, 2)
    )
    brainPlot[[i]] <- ggplot(data = plot_df, aes(x = ActualAge, y = PredictedAge)) +
        geom_point(color = "steelblue", size = 3, alpha = 0.8) +
        geom_smooth(method = "lm", color = "black", se = FALSE) +
        annotate("text",
            x = min(plot_df$ActualAge, na.rm = TRUE),
            y = max(plot_df$PredictedAge, na.rm = TRUE),
            label = annotation_text,
            hjust = 0,
            vjust = 1,
            size = 4
        ) +
        labs(
            title = "Oligodendrocytes",
            x = "Chronological Age",
            y = paste0("Predicted Age(", c("SC", "Pseudobulk", "Bootstrap")[i], ")")
        ) +
        theme_bw() +
        theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5, size = 12),
            axis.title = element_text(size = 14),
            axis.text = element_text(size = 12)
        )
}

ggpubr::ggarrange(plotlist = brainPlot, nrow = 1)
```

# Tutorial Example 5: Apply Pasta to Pbmc scRNA-seq dataset

```{r pasta, eval = FALSE}
# Load required libraries
library(OmniAgeR)
library(glmnet)
library(ggplot2)
library(patchwork)
library(ggpubr) # For simplified plot annotations

data("ScPbmcExample")
lognormMatrix <- as.matrix(SummarizedExperiment::assay(ScPbmcExample, "logcounts"))
cellMeta <- as.data.frame(SummarizedExperiment::colData(ScPbmcExample))
groups <- paste(cellMeta$donor_id, cellMeta$celltype, sep = "_")
uniqueGroups <- unique(groups)

 pseudobulkDataList <- lapply(uniqueGroups, function(grp) {
 rowSums(lognormMatrix[, groups == grp, drop = FALSE])
})

pseudobulkMatrix <- do.call(cbind, pseudobulkDataList)
colnames(pseudobulkMatrix) <- uniqueGroups

donorIdx <- match(uniqueGroups, groups)
donorAges <- cellMeta$age[donorIdx]
# Apply the PASTA clock
# filter_genes = TRUE: Selects only the genes required by the PASTA model.
# rank_norm = TRUE: Applies the rank-normalization required by PASTA.
pastaRes <- suppressMessages(pastaScores(pseudobulkMatrix, filterGenes = TRUE,
                        rankNorm = TRUE))
```

```{r pasta-plot, eval = FALSE, fig.width=4, fig.height=4, out.width="100%"}
# Prepare the data frame for plotting
plot_df <- data.frame(
    ActualAge = donorAges,
    Prediction = as.numeric(pastaRes$PASTA)
)

# Create the scatter plot
ggplot(data = plot_df, aes(x = ActualAge, y = Prediction)) +
    geom_point(color = "steelblue", size = 3, alpha = 0.8) +
    geom_smooth(method = "lm", color = "black", se = FALSE) +
    ggpubr::stat_cor(
        method = "pearson",
        label.x.npc = "left",
        label.y.npc = "top",
        hjust = 0,
        size = 5
    ) +
    labs(
        title = NULL,
        x = "Chronological Age",
        y = "PASTA Score"
    ) +
    theme_bw() +
    theme(
        plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.position = "none"
    )
```

# Tutorial Example 7: Calculate CPR and CHIP Scores

Here, we use DNAm data to calculate the CRP score and CHIP score.

```{r crp-chip, message = FALSE, warning = FALSE}
library(OmniAgeR)
data(dnamExample)
beta_matrix <- dnamExample[[1]]
crpRes <- compCRP(hannumBmiqM)
print(lapply(crpRes,head))

chipRes <- compCHIP(hannumBmiqM)
print(lapply(chipRes, head))
```


# Tutorial Example 8: Get freatures from Various Aging Biomarkers


```{r getMarkerWeights}
library(OmniAgeR)
# 1. Get probes for a specific chronological aging predictor
chronologicalProbes <- suppressMessages(getMarkerWeights(clockNames = "Horvath2013"))
print(head(chronologicalProbes$Horvath2013))
```


# Tutorial Example 9: Celltype fraction clock based on DNAm
We apply the DNAm Cell-type Fraction (CTF) Clock to 50 blood samples (EPIC array) from the TZH cohort to predict chronological age.

```{r dnam-ctf-clock, message = FALSE, warning = FALSE}
library(OmniAgeR)
library(randomForest)
library(ggplot2)
library(patchwork)
library(ggpubr)
library(SummarizedExperiment)
ctfExample <- loadOmniAgeRdata(
    "omniager_tzh_example_ctf",
    verbose = FALSE
)

phenoTypesTzh <- ctfExample[[1]]
ctfExampleSe <- SummarizedExperiment(
  assays = list(ctf=t(ctfExample[[2]]))
)

dnamCTFClockOut <- dnamCTFClock(x = ctfExampleSe)
```

```{r dnam-ctf-clock-plot, fig.width=4, fig.height=4, out.width="100%"}
plot_df <- data.frame(ActualAge = phenoTypesTzh$Age, PredictedAge = dnamCTFClockOut)
cor_test_result <- cor.test(plot_df$ActualAge, plot_df$PredictedAge)
correlation <- cor_test_result$estimate
p_value <- cor_test_result$p.value
mae <- mean(abs(plot_df$PredictedAge - plot_df$ActualAge))
p_value_formatted <- ifelse(p_value < 0.001,
    formatC(p_value, format = "e", digits = 2),
    round(p_value, 3)
)
annotation_text <- paste0(
    "R = ", round(correlation, 3), "\n",
    "P = ", p_value_formatted, "\n",
    "MAE = ", round(mae, 2)
)
ggplot(data = plot_df, aes(x = ActualAge, y = PredictedAge)) +
    geom_point(color = "steelblue", size = 3, alpha = 0.8) +
    geom_smooth(method = "lm", color = "black", se = FALSE) +
    annotate("text",
        x = min(plot_df$ActualAge, na.rm = TRUE),
        y = max(plot_df$PredictedAge, na.rm = TRUE),
        label = annotation_text,
        hjust = 0,
        vjust = 1, size = 5
    ) +
    labs(
        title = NULL,
        x = "Chronological Age",
        y = paste0("Predicted Age(DNAm_CTF)")
    ) +
    theme_bw() +
    theme(
        plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12)
    )


```

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

